---
title: "Report on Snickers and Coca-Cola Prices"
author: "Lagos Team"
date: "7 November 2021"
geometry: margin=2cm
fontsize: 9pt
output:
  html_document:
    df_print: paged
  pdf_document:
---

```{r echo=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
# Set graph size
#knitr::opts_chunk$set(echo = FALSE, out.width = "50%" )#fig.asp = 0.5, fig.width = 7, out.width = "90%" )

# Loading packages with pacman
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(tidyverse, modelsummary, viridis,
               kableExtra, patchwork)

# Import data
df <- read.csv( 'https://raw.githubusercontent.com/nszoni/lagos-team/main/coding-1/data/Data_DA1_HW1_Lagos.csv' )

```


```{r echo=FALSE, include=FALSE}

### Data cleaning


df_clean <- df %>%
  mutate(google_rating = as.numeric(ifelse(google_rating!='N/A', substr(google_rating,1,3), NA)), # 1. Correct ratings and NAs
         number_of_reviews = as.numeric(ifelse(number_of_reviews=='N/A', NA, number_of_reviews)), # 2. Correct number of reviews NA
         shop_type_group = as.factor(ifelse(shop_type %in% c('Grocery', 'Gas station', "Supermarket", "Other"), "Big shop", "Small shop")), # 3. Create group for shop size
         is_discounted = as.factor(ifelse(is_discounted=='Y', 1, 0)), # 4. Replace booelan flags
         has_website = as.factor(ifelse(has_website=='Y', 1, 0)),
         longitude = as.numeric(longitude), # 5. Data casting
         latitude = as.numeric(latitude),
         google_rating = as.numeric(google_rating),
         number_of_reviews = as.integer(number_of_reviews),
         collection_date = as.POSIXct(collection_date),
         district = as.factor(district),
         item_name = ifelse(item_name == 'Coca Cola, 0.5l plastic bottle', 'Coke 0.5l', 'Snickers 50g')) %>% #5. Shortening item names
  filter(is_discounted == 0)   # not discounted

```

## Introduction

Can we encounter lower prices in the outer districts then in the inner disticts? What if we go farther from the city center? In order to examine these questions, product prices were collected from two districts of Budapest, the 13th and the 10th, representing the inner and outer parts of the city. Our objective with the data collection was to compare retail prices of two chosen products, Coca-Cola 500ml plastic bottle and Snickers 50gr chocolate bar in the two areas. In order to examine the differences in the price distributions, product tags were gathered from shops of different sizes including supermarkets, gas stations, as well as small local groceries.
Our data set and artifacts can be found on our [Github repo](https://github.com/nszoni/lagos-team/)


## Data Quality and Cleaning
The focus of our analysis is on the prices of the two products. In order to increase the comparability between the registered prices, we recorded if the prducts were offered at a discounted price. Items with discounted price were later excluded from our analysis (2 records). Additional data wrangling was done in order to prepare our data for analysis: missing values were encoded with the conventional NA, product names were shortened and all the variables in the dataset were assigned proper data type. Moreover, we created two groups from the recorder shop types, were we aggregated the shops into big supermarket chains and small local groceries categories.


## Descriptive Statistics
To further understand and analyze the price differential for both the products in the two selected districts, a statistics summary table provided below was the first step. It includes the mean, median, standard deviation, minimum price, maximum price, range, 5th percentile and 95th percentile along with missing values. 


```{r, echo=FALSE, fig.align = 'center', }

#TODO: add kableExtra tabular form

  
# Transpose table
df_clean_wide <- df_clean %>%
  select(shop_name, item_name, district, price_huf) %>%
  spread(key = item_name, value = price_huf ) %>%
  arrange(district)
# Summary table
P95 <- function(x){quantile(x,0.95,na.rm=T)}
P05 <- function(x){quantile(x,0.05,na.rm=T)}
Range <- function(x){max(x, na.rm = TRUE) - min(x, na.rm = TRUE)}
Missing <- function(x){sum(is.na(x))}
datasummary((`District` = as.factor(district))  * ( (`Coke 0.5l` = `Coke 0.5l`  ) + (`Snickers 50g` = `Snickers 50g`  ) )  ~
              N +
              Missing +
              Mean * Arguments(fmt = "%.1f") +
              Median +
              SD * Arguments(fmt = "%.2f") +
              Min  +
              Max  +
              Range +
              P05  +
              P95 ,
              fmt = 0,
              data = df_clean_wide,
              title = 'Descriptive statistics of product prices by district',
              notes = '') %>%
    kableExtra::kable_styling(latex_options = "hold_position", font_size = 8)



```

As per the table, intuitively, the mean for Coke 0.5l bottle is slightly higher in the inner 13th district (HUF 318.4) as compared to out 10th district (HUF 304.5). While on the other hand, for Snickers 50grams there isn’t a huge difference in the mean price with 13th district having lower mean compared to 10th district  (district 13th = HUF 192.6 and district 10th = 193.8).
Also we can observed there is less variation in price of Snicker bars and more variation in the price of Coke. Furthermore, it was noticed that the range of prices for the both the products were around HUF 200 (in both the districts). Therefore, our team decided to further investigate this point and created two main question for analysis. For each product, the first question would highlight the price difference between districts and second to analyze price difference based on distance from the city center in both districts. 


## Price Distribution based on District
The team decided to select a combination of both histogram and density plot to show the distribution of prices for both the products in each district. Histogram with a bin size of 10 was used because the price difference and frequency price could be easily displayed, while density plot shows how the distribution looks like, making it easier to visualize. 

```{r echo = F, warning=FALSE, fig.height = 3, fig.align="center"}

# Histograms

# Coke distribution
hist_fig_1 <- df_clean %>% filter(item_name=='Coke 0.5l') %>% 
                ggplot() + 
                geom_histogram(aes(price_huf, y=..density.., fill=district), binwidth = 30, alpha=0.6, position="dodge", bins = 10) +
                geom_density(aes(price_huf, y=..density.., color=district), position="identity", alpha=0.5) +
                labs(x = "Price (HUF)",
                     y = "Density",
                     title = "Coca Cola 0.5l bottle"
                     # caption = "Max price was HUF 449 & HUF 439 in district 10 & 13, respectively"
                     ) +
                scale_fill_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                theme_classic()


#Snickers distribution
hist_fig_2 <- df_clean %>% filter(item_name=='Snickers 50g') %>% 
                ggplot() + 
                geom_histogram(aes(price_huf, y=..density.., fill=district), binwidth = 30, alpha=0.6, position="dodge", bins = 10) +
                geom_density(aes(price_huf, y=..density.., color=district), position="identity", alpha=0.5) +
               labs(x = "Price (HUF)",
                     y = "Density",
                     title = "Snickers 50g bar"
                     # caption = "Max price was HUF 319 & HUF 299 in district 10 & 13, respectively"
                    ) +
                scale_fill_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                theme_classic()

hist_fig_1 + hist_fig_2 + plot_layout(guides = "collect")

```

For Coke 0.5l, based on the graph below, majority of the prices (for both districts) is in the range of HUF 300 to HUF 320 which compliments the mean calculations. Overall, both the districts tends to have the same distribution with slightly right long tails. The distribution for Snickers 50grams is also similar in both districts i.e. slight right skewed. 


```{r echo=F, warning=F, message=F, fig.show='hide'}
# Boxplots


df_clean %>% 
  filter(item_name == 'Snickers 50g') %>% 
  ggplot( aes(x = factor(district), y = price_huf)) +
    stat_boxplot(aes(group = factor(district)), geom = "errorbar", width = 0.2, color = "black", size = 0.3) +
    geom_boxplot(aes(group = factor(district)),  color = viridis(2, begin=0.2, end=0.8), fill = viridis(2, begin=0.2, end=0.8), size = 0.4, width = 0.15, alpha = 0.4) + # , outlier.shape = NA
    # geom_jitter(aes(color = factor(district)), position=position_jitter(0.1), size = 1.5, show.legend=F) +
    labs(x = "District",y = "Price",
         subtitle = "of Coca Cola 0.5l bottle",
       title = "Price desnity by district (HUF)",
       caption = "lagos-team") +
    scale_y_continuous(limits = c(100,350), breaks = seq(0,350,50)) +
    scale_color_viridis(discrete = TRUE, option = "D", begin = 0.2, end=0.8)+
    theme_bw() 


df_clean %>% 
  filter(item_name == 'Coke 0.5l') %>% 
  ggplot(aes(x = factor(district), y = price_huf)) +
    stat_boxplot(aes(group = factor(district)), geom = "errorbar", width = 0.2, color = "black", size = 0.3) +
    geom_boxplot(aes(group = factor(district)),  color = viridis(2, begin=0.2, end=0.8), fill = viridis(2, begin=0.2, end=0.8), size = 0.4, width =  0.15, alpha = 0.4) +  # , outlier.shape = NA
    # geom_jitter(aes(color = factor(district)), position=position_jitter(0.1), size = 1.5, show.legend=F) +
    labs(x = "District",y = "Price",
          subtitle = "of Snickers 50g bar",
      title = "Price desnity by district (HUF)",
       caption = "lagos-team") +
    scale_y_continuous(limits = c(200, 450), breaks = seq(0,450,50)) +
    scale_color_viridis(discrete = TRUE, option = "D", begin = 0.2, end=0.8)+
    theme_bw() 

```


## Price Distribution based on Distance
To display the results of our second analysis, "relationship between price and distance from center" scatter plots were used to for both the product. 

```{r echo=F, fig.align='center', fig.height=3, warning=FALSE, message=FALSE}
# Scatterplots 

scatter_fig_1 <- df_clean %>% 
                    filter(item_name == 'Snickers 50g') %>% 
                    ggplot(aes(x = distance_from_city_centre_km, y = price_huf, color = factor(district))) +
                      geom_point( size = 3,  shape = 16, alpha = 0.9,  na.rm = TRUE) +
                      geom_smooth(method='lm',  linetype = "dashed", se = F) +
                      labs(x = "Distance from city centre (km)",y = "Price (HUF)",
                        title = "Coca Cola 0.5l bottle",
                        caption = "District 13 upward slop, District 10 downward slop") +
                      scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                      theme_bw()


scatter_fig_2 <- df_clean %>% 
                    filter(item_name == 'Coke 0.5l') %>% 
                    ggplot(aes(x = distance_from_city_centre_km, y = price_huf, color = factor(district))) +
                      geom_point( size = 3,  shape = 16, alpha = 0.9,  na.rm = TRUE) +
                      geom_smooth(method='lm', linetype = "dashed", se = F) +
                      labs(x = "Distance from city centre (km)",y = "Price (HUF)",
                        title = "Snickers 50g bar",
                        caption = "District 13 upward slop, District 10 downward slop") +
                      scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                      theme_bw()

scatter_fig_1 + scatter_fig_2 + plot_layout(guides = "collect")
```

As shown in the graph below, it can be observed that for both products in District 13 the price increased as you moved away from the city center, Thus an upwards slop of the line of best fit. In contrary to that, in District 10 prices for both the products decreased as you go further away from the city center which is shown by downward slop of the line of best fit . Adding to that, both the graphs further show a unique crossover of prices in the middle which is between the range of 4kms to 6kms away from city center and highlights that prices tend to high in that region regardless of the product.  


# Conclusion
Segueing on the analysis, the district and distance from city center in that district are trivial to understand the distribution of price change for both the products reviewed. It is not wrong to conclude that on average the price is similar in both the districts, but customers buying any of these products in the range of  4km – 6km distance from the city center, regardless of the district, tend to pay more as compared other areas in the district. To have a more conclusive understanding of this distribution, more factors and variables could be added to the analysis.  
