---
title: "Report on Snickers and Coca-Cola Prices"
author: "Lagos Team"
date: "7 November 2021"
geometry: margin=1.8cm
fontsize: 9pt
output:
  pdf_document:
  html_document:
    df_print: paged
header-includes: |
  \usepackage{titling}
  \setlength{\droptitle}{-5em} 
---

```{r echo=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
# Set graph size
#knitr::opts_chunk$set(echo = FALSE, out.width = "50%" )#fig.asp = 0.5, fig.width = 7, out.width = "90%" )

# Loading packages with pacman
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(tidyverse, modelsummary, viridis,
               kableExtra, patchwork)

# Import data
df <- read.csv( 'https://raw.githubusercontent.com/nszoni/lagos-team/main/coding-1/data/Data_DA1_HW1_Lagos.csv' )

```

```{r echo=FALSE, include=FALSE}

### Data cleaning


df_clean <- df %>%
  mutate(google_rating = as.numeric(ifelse(google_rating!='N/A', substr(google_rating,1,3), NA)), # 1. Correct ratings and NAs
         number_of_reviews = as.numeric(ifelse(number_of_reviews=='N/A', NA, number_of_reviews)), # 2. Correct number of reviews NA
         shop_type_group = as.factor(ifelse(shop_type %in% c('Grocery', 'Gas station', "Supermarket", "Other"), "Big shop", "Small shop")), # 3. Create group for shop size
         is_discounted = as.factor(ifelse(is_discounted=='Y', 1, 0)), # 4. Replace booelan flags
         has_website = as.factor(ifelse(has_website=='Y', 1, 0)),
         longitude = as.numeric(longitude), # 5. Data casting
         latitude = as.numeric(latitude),
         google_rating = as.numeric(google_rating),
         number_of_reviews = as.integer(number_of_reviews),
         collection_date = as.POSIXct(collection_date),
         district = as.factor(district),
         item_name = ifelse(item_name == 'Coca Cola, 0.5l plastic bottle', 'Coke 0.5l', 'Snickers 50g')) %>% #5. Shortening item names
  filter(is_discounted == 0)   # not discounted

```

## Introduction

Can we encounter lower prices in the outer districts then in the inner districts? What if we go farther away from the city center? In order to examine these questions, product prices were collected from two districts of Budapest, the $13^{th}$ and the $10^{th}$, representing the inner and outer parts of the city. Our objective with the data collection was to compare retail prices of two chosen products, Coca-Cola 500ml plastic bottle and Snickers 50gr chocolate bar in the two areas. In order to examine the differences in the price distributions, product tags were gathered from shops of different sizes including supermarkets, gas stations, as well as small local groceries.
Our data set and artifacts can be found on our [Github repo](https://github.com/nszoni/lagos-team/tree/main/coding-1)


## Data Quality and Cleaning
The focus of our analysis is on the prices of the two products. In order to increase the comparability between the registered prices, we recorded if the products were offered at a discounted price. Items with discounted price were later excluded from our analysis (2 records). Additional data wrangling was done in order to prepare our data for analysis: missing values were encoded with the conventional NA, product names were shortened and all the variables in the dataset were assigned proper data type. Moreover, we created two groups from the shop types, where we aggregated the shops into big supermarket chains and small local groceries categories.


## Descriptive Statistics
To further understand and analyze the price differences between the the two selected districts, a summary statistics table was created.


```{r, echo=FALSE, fig.align = 'center', }

#TODO: add kableExtra tabular form

  
# Transpose table
df_clean_wide <- df_clean %>%
  select(shop_name, item_name, district, price_huf) %>%
  spread(key = item_name, value = price_huf ) %>%
  arrange(district)
# Summary table
P95 <- function(x){quantile(x,0.95,na.rm=T)}
P05 <- function(x){quantile(x,0.05,na.rm=T)}
Range <- function(x){max(x, na.rm = TRUE) - min(x, na.rm = TRUE)}
Missing <- function(x){sum(is.na(x))}
datasummary((`District` = as.factor(district))  * ( (`Coke 0.5l` = `Coke 0.5l`  ) + (`Snickers 50g` = `Snickers 50g`  ) )  ~
              N +
              Missing +
              Mean * Arguments(fmt = "%.1f") +
              Median +
              SD * Arguments(fmt = "%.2f") +
              Min  +
              Max  +
              Range +
              P05  +
              P95 ,
              fmt = 0,
              data = df_clean_wide,
              title = 'Descriptive statistics of product prices by district',
              notes = '') %>%
    kableExtra::kable_styling(latex_options = "hold_position", font_size = 8)



```

As per the summary statistics, the mean price of Coke is slightly higher in the inner $13^{th}$ district (HUF 318.4) compared to the $10^{th}$ district (HUF 304.5). As for Snickers, we cannot see considerable difference between average prices (HUF 1.2).
Regarding the spread of prices, there's less variation in Snickers' prices compared to Coke's. The range of prices for both the products were in the interval of 180 to 219, while 95% of the Coke prices were lower than 380 and 426 for the 10th and 13th district, respectively.

## Price Distribution based on District
We chose histograms to show the distribution of prices for both the products in each district. We set the bin size to HUF 50 and the positioning of bars to 'dodged', as it is easier to interpret a HUF 50 step in prices, while it did not impose additional clutter.

Overall, Coke prices have distribution with long right tails in both districts, while in case of Snickers, the distribution is centered around HUF 200.The majority of the shops offer Coke in the price interval between HUF 275 and 325 in both district while in case of Snickers the mode is at the 175-225 price bin.

```{r echo = F, warning=FALSE, fig.height = 3, fig.align="center"}

# Histograms

# Coke distribution
hist_fig_1 <- df_clean %>% filter(item_name=='Coke 0.5l') %>% 
                ggplot() + 
                geom_histogram(aes(price_huf, y=..count../sum(count), fill=district), binwidth = 50, alpha=0.8, position="dodge") +
                labs(x = "Price (HUF)",
                     y = "Density",
                     title = "Coca Cola 0.5l bottle"
                     # caption = "Max price was HUF 449 & HUF 439 in district 10 & 13, respectively"
                     ) +
                scale_fill_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                theme_classic() +
                scale_y_continuous(labels = scales::percent_format(accuracy = 5L))


#Snickers distribution
hist_fig_2 <- df_clean %>% filter(item_name=='Snickers 50g') %>% 
                ggplot() + 
                geom_histogram(aes(price_huf, y=..count../sum(count), fill=district), binwidth = 50, alpha=0.8, position="dodge") +
               labs(x = "Price (HUF)",
                     y = "Density",
                     title = "Snickers 50g bar"
                     # caption = "Max price was HUF 319 & HUF 299 in district 10 & 13, respectively"
                    ) +
                scale_fill_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                theme_classic() +
                scale_y_continuous(labels = scales::percent_format(accuracy = 5L))

hist_fig_1 + hist_fig_2 + plot_layout(guides = "collect")

```

*Observations with extreme price values at the tail of the distribution are driven by the prices of gas stations.*

```{r echo=F, warning=F, message=F, fig.show='hide'}
# Boxplots


df_clean %>% 
  filter(item_name == 'Snickers 50g') %>% 
  ggplot( aes(x = factor(district), y = price_huf)) +
    stat_boxplot(aes(group = factor(district)), geom = "errorbar", width = 0.2, color = "black", size = 0.3) +
    geom_boxplot(aes(group = factor(district)),  color = viridis(2, begin=0.2, end=0.8), fill = viridis(2, begin=0.2, end=0.8), size = 0.4, width = 0.15, alpha = 0.4) + # , outlier.shape = NA
    # geom_jitter(aes(color = factor(district)), position=position_jitter(0.1), size = 1.5, show.legend=F) +
    labs(x = "District",y = "Price",
         subtitle = "of Coca Cola 0.5l bottle",
       title = "Price desnity by district (HUF)",
       caption = "lagos-team") +
    scale_y_continuous(limits = c(100,350), breaks = seq(0,350,50)) +
    scale_color_viridis(discrete = TRUE, option = "D", begin = 0.2, end=0.8)+
    theme_bw() 


df_clean %>% 
  filter(item_name == 'Coke 0.5l') %>% 
  ggplot(aes(x = factor(district), y = price_huf)) +
    stat_boxplot(aes(group = factor(district)), geom = "errorbar", width = 0.2, color = "black", size = 0.3) +
    geom_boxplot(aes(group = factor(district)),  color = viridis(2, begin=0.2, end=0.8), fill = viridis(2, begin=0.2, end=0.8), size = 0.4, width =  0.15, alpha = 0.4) +  # , outlier.shape = NA
    # geom_jitter(aes(color = factor(district)), position=position_jitter(0.1), size = 1.5, show.legend=F) +
    labs(x = "District",y = "Price",
          subtitle = "of Snickers 50g bar",
      title = "Price desnity by district (HUF)",
       caption = "lagos-team") +
    scale_y_continuous(limits = c(200, 450), breaks = seq(0,450,50)) +
    scale_color_viridis(discrete = TRUE, option = "D", begin = 0.2, end=0.8)+
    theme_bw() 

```

## Price Distribution based on Distance
We used scatterplots to display our findings on the relationship between the distance from city center and the product prices.

As shown in the graph below, for both products there is a positive association between the distance and the price in the $13^{th}$ district. In contrary to that, $10^{th}$ district prices tend to decrease with distance.

```{r echo=F, fig.align='center', fig.height=3, warning=FALSE, message=FALSE}
# Scatterplots 

scatter_fig_1 <- df_clean %>% 
                    filter(item_name == 'Snickers 50g') %>% 
                    ggplot(aes(x = distance_from_city_centre_km, y = price_huf, color = factor(district))) +
                      geom_point( size = 3,  shape = 16, alpha = 0.9,  na.rm = TRUE) +
                      geom_smooth(method='lm',  linetype = "dashed", se = F) +
                      labs(x = "Distance from city centre (km)",y = "Price (HUF)",
                        title = "Coca Cola 0.5l bottle") +
                      scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                      theme_bw()


scatter_fig_2 <- df_clean %>% 
                    filter(item_name == 'Coke 0.5l') %>% 
                    ggplot(aes(x = distance_from_city_centre_km, y = price_huf, color = factor(district))) +
                      geom_point( size = 3,  shape = 16, alpha = 0.9,  na.rm = TRUE) +
                      geom_smooth(method='lm', linetype = "dashed", se = F) +
                      labs(x = "Distance from city centre (km)",y = "Price (HUF)",
                        title = "Snickers 50g bar") +
                      scale_colour_viridis_d( begin = 0.2, end = 0.8, name = "District") +
                      theme_bw()

scatter_fig_1 + scatter_fig_2 + plot_layout(guides = "collect")
```

## Conclusion
To sum up, our analysis indicates that prices in the inner district are generally higher than the outer districts of Budapest. After examining the relationship between distance and price, we can conclude that there exists an upward tendency in the $13^{th}$ district, as opposed to the $10^{th}$ district where the relationship seems negative.

**_NOTE:_** There are two further analysis (box whisker plot & t-test) made as 
a part of this exercise in R Markdown but included in the report.
